Cudele: A File System with Programmable Consistency and Durability

### 2017-03-29 paper/sections/0_abstract.tex
paper: wordsmith and push related work backwards
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

HPC file systems are abandoning POSIX because the synchronization and
serialization overheads are too costly -- and often uneccessary -- for their
applications.  We provide an API for the client to (1) merge updates into the
global namespace and (2) assign policies to subtrees so that other clients
cannot interfere with the decoupled namespace.  This allows administrators to
optimize subtrees within the same namespace for different workloads. We draw
conclusions about the performance impact of previously unexplored
consistency/durability metadata designs and show that consistency can cause a
104\(\times\) slow down while merging updates (7\(\times\) slow down) and
maintaining durability (\(10\times\) slow down) have a more reasonable cost.

>>>>

HPC file systems are abandoning POSIX because the synchronization and
serialization overheads are too costly -- and often uneccessary -- for their
applications.  We provide an API for the client to (1) decouple subtrees from
the global namespace, (2) merge metadata updates back into the global namespace
and (3) assign policies to subtrees so that other clients cannot interfere with
the decoupled namespace.  This allows administrators to optimize subtrees
within the same namespace for different workloads. We draw conclusions about
the performance impact of previously unexplored consistency/durability metadata
designs and show that consistency can cause a 104\(\times\) slow down while
merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down) have a more reasonable cost.

### 2017-03-29 paper/sections/1_introduction.tex
paper: wordsmith and push related work backwards
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% What is the problem?
Today's client-server based file system metadata services have scalability
problems. It takes a lot of resources to service POSIX metadata requests so
applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. Unfortunately,
provisioning a metadata server for every client is expensive and complicated.

% How does this fit in current landscape?
Current hardware evolution and the rise of software-defined storage storage,
which uses techniqes like erasure coding, replication, and partitioning, have
ushered in a new era of HPC computing; architectures are transitioning from
complex storage stacks with burst buffer, file system, object store, and tape
tiers to a two layer stack with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}. This trend exacerbates the metadata
scalability problem

% What is HPC doing?
To address this trend, HPC research has advocated client side processing and an
emphasis on serverless metadata services.  The techniques propose new metadata
management techniques that reduce synchronization and serialization overheads
by transferring these responsibilities to the client. We call this approach
``decoupling the namespace" and the semantics of consistency differ amongst
systems.  While the performance benefits are obvious for these users,
applications that rely on stronger consistency or durability guarantees must be
re-written or deployed on a different system.

@@

% What did we do
We propose subtree policies, an interface that lets future programmers control
how the system manages different parts of the namespace.  For performance one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  We
present Cudele, a prototype programmable file system that supports different
degrees of consistency and durability by exposing mechanisms used within the
file system as a client library.  Cudele supports 3 forms of consistency
(invisible, eventual, and strong) and 3 degrees of durability (none, local, and
global) giving the administrator a wide range of policies and optimizations
that can be custom fit to an application. Our contributions:

@@

  \item an API for assigning and programming policies to subtrees in the file
  system namespace.

  \item an analysis of recently proposed research systems compared against
  previously unexplored metadata designs.

>>>>

% What is the problem?
Today's client-server based file system metadata services have scalability
problems. It takes a lot of resources to service POSIX metadata requests so
applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
metadata server for every client is expensive and complicated. This problem is
exacerbated by current trends in HPC.  Architectures are transitioning from
complex storage stacks with burst buffer, file system, object store, and tape
tiers to a more simplified stack with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
instead of data movement.

% What is HPC doing?
HPC's unique requirements for file systems ({\it e.g.}, fast creates) and well
defined workloads ({\it e.g.}, workloads) make relaxing POSIX sensible. Without
the constraints of strong consistency and durability, HPC systems can now do
more client side processing with an emphasis on serverless metadata services.
One popular approach is to ``decouple the namespace", where clients lock the
subtree they want exclusive access to so other clients cannot interfere. This
delayed merge ({\it i.e.} a form of eventual consistency) and relaxed
durability improves performance and scalability by avoiding the costs of RPCs,
synchronization, false sharing, and serialization.  While the performance
benefits are obvious for these users, applications that rely on stronger
consistency or durability guarantees must be re-written or deployed on a
different system. The consistency and durability semantics for these systems is
shown in Table~\ref{table:namespaces}.

@@

\begin{table}
\begin{tabular}{ r | l | l }
              & Decoupled & Global    \\
              & Namespace & Namespace \\\hline
  Example     & BatchFS~\cite{zheng:pdsw2014-batchfs} & CephFS~\cite{weil:sc2004-dyn-metadata} \\
              & DeltaFS~\cite{zheng:pdsw2015-deltafs} & IndexFS~\cite{ren:sc2014-indexfs}      \\
  Consistency & eventual & strong     \\
  Durability  & node local & global  \\
\end{tabular}
\caption{State-of-the-art systems in HPC improve file system metadata
performance by relaxing consistency and durability
guarantees.\label{table:namespaces}}
\end{table}

% What did we do
We propose subtree policies, an interface that lets future programmers control
how the storage system manages different parts of the file system namespace.  For performance one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  We
present Cudele, a prototype programmable file system that supports different
degrees of consistency and durability by exposing mechanisms used within the
file system as a client library.  Cudele supports 3 forms of consistency
(invisible, eventual, and strong) and 3 degrees of durability (none, local, and
global) giving the administrator a wide range of policies and optimizations
that can be custom fit to an application. Our contributions:

@@

  \item an API for programming consistency/durability policies and assigning
  them to subtrees in the file system namespace.

  \item a comparison of the strategies used in recently proposed research systems against
  previously unexplored metadata designs.

### 2017-03-29 paper/sections/2_related_work.tex
paper: cleanup terminology in section 4
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: wordsmith and push related work backwards
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% approaches to load balancing (strong = more resuorces per unit work, weak =
% fixed resource per work unit)
One approach for improving metadata performance and scalability is to alleviate
overloaded servers by load balancing metadata IO across a cluster. Common
techniques include partitioning metadata when there are many writes and
replicating metadata when there are many reads. For example, IndexFS partitions
directories and clients write to different partitions by grabbing leases and
caching ancestor metadata for path traversal; it does well for strong scaling
because servers can keep more inodes in the cache which results in less RPCs.
Alternatively, ShardFS replicates directory state so servers do not need to
contact peers for path traversal; it does well for read workloads because all
file operations only require 1 RPC and for weak scaling because requests will
never incurr extra RPCs due to a full cache.  CephFS employs both techniques to
a lesser extent; directories can be replicated or sharded but the caching and
replication policies do not change depending on the balancing technique.
Despite the performance benefits these techniques add complexity and jeopardize
the robustness and performance characteristics of the metadata service because
the systems now need (1) policies to guide the migration decisions and (2)
mechanisms to address inconsistent states across servers.

% policies
Setting policies for migrations is arguably more difficult than adding the
migration mechanisms themselves.  For example, IndexFS and CephFS use the
GIGA+~\cite{patil:fast2011-giga} technique for partitioning directories at a predefined
threshold and using lazy synchronization to redirect queries to the server that
``owns" the targeted metadata.  Determining when to partition directories and
when to migrate the directory fragments are policies that vary between systems:
GIGA+ partitions directories when the size reaches a certain number of files
and migrates directory fragments immediately; CephFS partitions directories
when they reach a threshold size or when the write temperature reaches a
certain value and migrates directory fragments when the hosting server has more
load than the other servers in the metadata cluster. Another policy is when and
how to replicate directory state; ShardFS replicates immediately and
pessimistically while CephFS replicates only when the read temperature reaches
a threshold.  There is a wide range of policies and it is difficult to address
with tunables and hard-coded design decisions.

% addressing inconsistency
In addition to the policies, distributing metadata across a cluster requiress
distributed transactions and cache coherence protocols to ensure strong
consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
directory state and uses optimistic concurrency control for conflicts; namely
it does the operation and if there is a conflict at verification time it falls
back to two-phase locking.  Another example is IndexFS's inode cache which
reduces RPCs by caching ancestor paths -- the locality of this cache can be
thrashed by random reads but performs well for metadata writes. For
consistency, writes to directories in IndexFS block until the lease expires
while writes to direcotries in ShardFS are slow for everyone as it either
requires serialization or locking with many servers; reads in IndexFS are
subject to cache locality while reads in ShardFS always resolve to 1 RPC.
Another example of the overheads of addressing inconsistency is how CephFS
maintains client sessions and inode caches for capabilities (which in turn make
metadata access faster). When metadata is exchanged between metadata servers
these sessions/caches must be flushed and new statistics exchanged with a
scatter-gather process; this halts updates on the directories and blocks until
the authoratitive metadata server responds.  These protocols are discusssed in
more detail in the next section but their inclusion here is a testament to the
complexity of migrating metadata.

@@

%IndexFS aggressively caches paths and
%handles permissions by handing out leases for metadata writes; metadata may
%only be modified when all leases have expired. 
%IndexFS~\cite{ren:sc2014-indexfs} aggressively caches pathnames and their
%permissions on the client servers. Modifications to metadata cached by clients
%is delayed until all client leases have expired. This reduces the RPC
%amplification to 1 when mutatating directory metadata ({\it e.g.,}
%\texttt{mkdir}, \texttt{chmod}, etc.) because clients are not querying MDSs for
%path traversals. The disadvantage of this approach is the high latency of
%mutation operations (reads to filenames).  ShardFS~\cite{xiao:socc2015-shardfs}
%replicates metadata (specifically directory lookup state) across the MDS
%cluster, reducing the RPC amplification to 1 for file operations ({e.g.,}
%\texttt{stat}, \texttt{chmod}, \texttt{chown}, etc.). Modifications to the
%directory lookup state are done with optimistic concurrecny control and fall
%back to retry if verification fails. The disadvantage of this appraoch is the
%high number of RPCs for maintaining directory metadata mutations (writes to
%directories).  CephFS also maintains an inode cache but its notion of leases
%are much shorter, on the order of microseconds.

% Non POSIX
High performance computing has unique requirements for file systems ({e.g.},
fast creates) and well-defined workloads (e.g., workflows) that make relaxing
POSIX sensible.  One popular approach is to allow clients to ``lock'' parts of
the namespace to improve performance and scalability by avoiding
synchronization, false sharing, and serialization.  BatchFS assumes the
application coordinates accesses to the namespace, so the clients can batch
local operations and merge with a global namespace image lazily. Similarily,
DeltaFS eliminates RPC traffic using subtree snapshots for non-conflicting
workloads and middleware for conflicting workloads. MarFS gives administrators
the ability to (1) lock ``project directories" and (2) allocate GPFS clusters
for demanding directory workloads. TwoTiers eliminates high-latencies by
storing metadata in a flash tier; apps lock the namespace so that metadata can
be accessed more quickly.  Unfortunately, decoupling the namespaces has costs:
(1) merging metadata state back into the global namespace is slow; (2) failures
are local to the failing node; and (3) the systems are not backwards
compatible.

@@

Decoupling the namespace delays metadata consistency and sacrifices durability. 
As shown in Table~\ref{table:namespaces}, metadata consistency is
provided by capabilities and metadata durability is addressed with a journal.
\begin{table}
\begin{tabular}{ r | l | l }
              & Decoupled & Global    \\
              & Namespace & Namespace \\\hline
  Example     & BatchFS~\cite{zheng:pdsw2014-batchfs} & CephFS~\cite{weil:sc2004-dyn-metadata} \\
              & DeltaFS~\cite{zheng:pdsw2015-deltafs} & IndexFS~\cite{ren:sc2014-indexfs}      \\
  Consistency & eventual & strong     \\
  Durability  & node local & journal  \\
\end{tabular}
\caption{State-of-the-art systems in HPC improve file system metadata
performance by relaxing consistency and durability
guarantees.\label{table:namespaces}}
\end{table}

>>>>

% approaches to load balancing (strong = more resuorces per unit work, weak =
% fixed resource per work unit)
One approach for improving metadata performance and scalability is to alleviate
overloaded servers by load balancing metadata IO across a cluster. Common
techniques include partitioning metadata when there are many writes and
replicating metadata when there are many reads. For example, IndexFS~\cite{ren:sc2014-indexfs} partitions
directories and clients write to different partitions by grabbing leases and
caching ancestor metadata for path traversal; it does well for strong scaling
because servers can keep more inodes in the cache which results in less RPCs.
Alternatively, ShardFS replicates directory state so servers do not need to
contact peers for path traversal; it does well for read workloads because all
file operations only require 1 RPC and for weak scaling because requests will
never incurr extra RPCs due to a full cache.  CephFS employs both techniques to
a lesser extent; directories can be replicated or sharded but the caching and
replication policies do not change depending on the balancing technique.
Despite the performance benefits these techniques add complexity and jeopardize
the robustness and performance characteristics of the metadata service because
the systems now need (1) policies to guide the migration decisions and (2)
mechanisms to address inconsistent states across servers.

% policies
Setting policies for migrations is arguably more difficult than adding the
migration mechanisms themselves.  For example, IndexFS and CephFS use the
GIGA+~\cite{patil:fast2011-giga} technique for partitioning directories at a predefined
threshold and using lazy synchronization to redirect queries to the server that
``owns" the targeted metadata.  Determining when to partition directories and
when to migrate the directory fragments are policies that vary between systems:
GIGA+ partitions directories when the size reaches a certain number of files
and migrates directory fragments immediately; CephFS partitions directories
when they reach a threshold size or when the write temperature reaches a
certain value and migrates directory fragments when the hosting server has more
load than the other servers in the metadata cluster. Another policy is when and
how to replicate directory state; ShardFS replicates immediately and
pessimistically while CephFS replicates only when the read temperature reaches
a threshold.  There is a wide range of policies and it is difficult to maneuver
tunables and hard-coded design decisions.

% addressing inconsistency
In addition to the policies, distributing metadata across a cluster requiress
distributed transactions and cache coherence protocols to ensure strong
consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
directory state and uses optimistic concurrency control for conflicts; namely
it does the operation and if there is a conflict at verification time it falls
back to two-phase locking.  Another example is IndexFS's inode cache which
reduces RPCs by caching ancestor paths -- the locality of this cache can be
thrashed by random reads but performs well for metadata writes. For
consistency, writes to directories in IndexFS block until the lease expires
while writes to directories in ShardFS are slow for everyone as it either
requires serialization or locking with many servers; reads in IndexFS are
subject to cache locality while reads in ShardFS always resolve to 1 RPC.
Another example of the overheads of addressing inconsistency is how CephFS
maintains client sessions and inode caches for capabilities (which in turn make
metadata access faster). When metadata is exchanged between metadata servers
these sessions/caches must be flushed and new statistics exchanged with a
scatter-gather process; this halts updates on the directories and blocks until
the authoratitive metadata server responds.  These protocols are discusssed in
more detail in the next section but their inclusion here is a testament to the
complexity of migrating metadata.

@@

%IndexFS aggressively caches paths and
%handles permissions by handing out leases for metadata writes; metadata may
%only be modified when all leases have expired. 
%IndexFS~\cite{ren:sc2014-indexfs} aggressively caches pathnames and their
%permissions on the client servers. Modifications to metadata cached by clients
%is delayed until all client leases have expired. This reduces the RPC
%amplification to 1 when mutatating directory metadata ({\it e.g.,}
%\texttt{mkdir}, \texttt{chmod}, etc.) because clients are not querying metadata servers for
%path traversals. The disadvantage of this approach is the high latency of
%mutation operations (reads to filenames).  ShardFS~\cite{xiao:socc2015-shardfs}
%replicates metadata (specifically directory lookup state) across the metadata server
%cluster, reducing the RPC amplification to 1 for file operations ({e.g.,}
%\texttt{stat}, \texttt{chmod}, \texttt{chown}, etc.). Modifications to the
%directory lookup state are done with optimistic concurrecny control and fall
%back to retry if verification fails. The disadvantage of this appraoch is the
%high number of RPCs for maintaining directory metadata mutations (writes to
%directories).  CephFS also maintains an inode cache but its notion of leases
%are much shorter, on the order of microseconds.

% Non POSIX
High performance computing has unique requirements for file systems ({\it
e.g.}, fast creates) and well-defined workloads ({\it e.g.}, workflows) that
make relaxing POSIX sensible.  BatchFS assumes the application coordinates
accesses to the namespace, so the clients can batch local operations and merge
with a global namespace image lazily. Similarily, DeltaFS eliminates RPC
traffic using subtree snapshots for non-conflicting workloads and middleware
for conflicting workloads. MarFS gives administrators the ability to lock
``project directories" and allocate GPFS clusters for demanding metadata
workloads. TwoTiers eliminates high-latencies by storing metadata in a flash
tier; apps lock the namespace so that metadata can be accessed more quickly.
Unfortunately, decoupling the namespace has costs: (1) merging metadata state
back into the global namespace is slow; (2) failures are local to the failing
node; and (3) the systems are not backwards compatible.

@@
